Red Wine Quality Analysis by Karun Gahlawat
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
#packageVersion('ggplot2')
library(knitr)
library(dplyr)
library(reshape2)
library(scales)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
rwdata <- read.csv("wineQualityReds.csv", header = TRUE) 
#### check data is ready
if(sum(complete.cases(rwdata)) == 1599){
  print("data complete")
} else {
  print("data incomplete please clean")
}
```
> **Brief Summary of Data**:

__Red Wine__ data consists of about 1500 with the following features

   1 - fixed acidity: most acids involved with wine or fixed or nonvolatile (do not evaporate readily)

   2 - volatile acidity: the amount of acetic acid in wine, which at too high of levels can lead to an unpleasant, vinegar taste

   3 - citric acid: found in small quantities, citric acid can add 'freshness' and flavor to wines

   4 - residual sugar: the amount of sugar remaining after fermentation stops, it's rare to find wines with less than 1 gram/liter and wines with greater than 45 grams/liter are considered sweet

   5 - chlorides: the amount of salt in the wine

   6 - free sulfur dioxide: the free form of SO2 exists in equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial growth and the oxidation of wine

   7 - total sulfur dioxide: amount of free and bound forms of S02; in low concentrations, SO2 is mostly undetectable in wine, but at free SO2 concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine

   8 - density: the density of water is close to that of water depending on the percent alcohol and sugar content

   9 - pH: describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale

   10 - sulphates: a wine additive which can contribute to sulfur dioxide gas (S02) levels, wich acts as an antimicrobial and antioxidant

   11 - alcohol: the percent alcohol content of the wine
   Output variable (based on sensory data): 

   12 - quality (score between 0 and 10)
   
   * We will explore this data with each feature and will try to explain what features contribute to a good rating.
   

# Univariate Plots Section


```{r echo=FALSE}
col_names <- colnames(rwdata)
for(c in col_names){
  if(c != "X"){
    print(c)
    pallete <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
    coli <- sample(1:length(pallete), 1, replace = TRUE)
    hists <- ggplot(rwdata, aes(x= rwdata[[c]])) + geom_histogram(col= pallete[coli]) + xlab(c)
    print(hists)
    scatters = ggplot(rwdata, aes(x= X, y= rwdata[[c]]), col= pallete(coli)) + geom_point(col= pallete[coli]) + ylab(c)
    print(scatters)  
  }
  
}

```






# Univariate Analysis

> fixed.acidity is positively skewed near normally distributed with large right tail. The majority of reading center around 6-8 with some values as far as 16. 

> volatile.acidity is centered around 0.4 and 0.6. Some outliers show as far as 1.5.

> citric.acid is not centered around a single value but spread over a broad range. Largest counts near 0 and 0.5. Some outliers around 1. Wow! 1? Might as well drink straight from a lime!!!

> residual.sugar is centered around 2 with some outliers as far as 8-16. Most observations are in the range 1-3.

> chlorides more than half of observations fall within 0.03 - 0.04. Outliers go as far as 0.6.

> free.sulfur.dioxide majority of observations lie within 0-20. They are un evenly spread out and are not centered around one value. outliers go as far as 80.

> total.sulfur.dioxide majority of observations lie within 0-50 and are not centered. Outliers go as far as 300!. Wow that is a lot of variation! Wonder if these high values should be ignored in analysis?

> density looks centered aroun .9975 with values almost symetrical both below and above this value but most observations near this value. Interesting... wonder if this factor is most responsible for quality given its symetry?

> pH like density this one is also centered and almost symetrical. Although it seem to have broader outliers than density. Wonder if this factor as well 
influences quality more than others? Or whether this factor has to be around certain value (3.25 in this case) to be drinkable? Should be interesting to find out!

> sulphates is centered around 0.5- 0.6 but are more spread out to the right (higher values) with some outliers as far as 2. Wonder wines with higher quality have low or high sulpfur content?

> alcohol highest count ~ 250 around 9.5 and the rest spread out between 9 and 12. Some outliers go as fas as 16! Usually, the wines I like have alcohol content between 12 and 15. Should be interesting to see what level dominates among high quality wines.

> quality ratings are in this order 5, 6, 7, 4, 8 and 3. Most observation fall in rating 5, then in rating 6, then in rating 7 and so on. We see none at far ends (0, 10). It is interesting to see how many fall in the mid range.


### What is the structure of your dataset?

> data set is set of ~ 1600 observations with 11 features. It is a two dimensional table with last column as the output column that classify each observation into 10 possible categories. Data is tidy as in all rows are observations and each column is a feature. No duplicates and no missing or NAs. It appears that this data is almost ready for training.

### What is/are the main feature(s) of interest in your dataset?

> Values are not scaled. Some features are centered around their mean values but most of them are spread out. This indicates towards non linear relationship with the final variable and among other factors as well. This also hints towards difficulty to find a valid relation ship among factors. Some interesting factors are density, pH, citric.acid, 

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

> I compared means of features with means of features where qaulity is 8 or greater. It gave some interesting insight on this dataset that most features have mean close to the quality means. There are a few features that stand a bit apart which could indicate some importance towards quality of wine.

### Did you create any new variables from existing variables in the dataset?

> I compared means of features with means of features where qaulity is 8 or greater. It gave some interesting insight on this dataset that most features have mean close to the quality means. There are a few features that stand a bit apart which could indicate some importance towards quality of wine.

### Did you create any new variables from existing variables in the dataset?

> No. There are plenty of variables to explore this data set of size 1600 rows.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

> I did not tidy any further this data base although I did filter some things out to explain some factors.


# Bivariate Plots Section



```{r echo=FALSE, Bivariate_Plots_Means}
means <- as.data.frame(sapply(rwdata, mean))
means_x = as.data.frame(means[2:nrow(means),1:ncol(means)])
colnames(means_x) <- c("Mean")
rownames(means_x) <- rownames(means)[2:nrow(means)]
means_quality <- filter(rwdata, quality >= 8)
means_quality <- as.data.frame(sapply(means_quality, FUN=mean))
means_quality <- as.data.frame(means_quality[2:nrow(means_quality),1:ncol(means_quality)])
colnames(means_quality) <- c("MeanQuality")
rownames(means_quality) <- rownames(means_x)
means_quality$Mean <- means_x$Mean 
means_quality
attach(means_quality)
m <- ggplot(means_quality, aes(x=rownames(means_quality))) + geom_point(aes(y=Mean, color="b"), show.legend=TRUE) + geom_point(aes(y=MeanQuality, color="r"), show.legend=TRUE) + ylab("Means") + xlab("Features") + scale_colour_manual(name="Means Comparison", values= c("b"="blue", "r"="red"), labels = c("Mean", "Mean of Qaulity >= 8"))
m <- m + theme(axis.text = element_text(angle=45, hjust = 1))
print(m)
pairs(rwdata[,2:13])

```

```{r echo=FALSE, Bivariate_Plots_Fixed.Acidity}
fixed.acidity_quality <- ggplot(rwdata, aes(x=quality, y=rwdata$fixed.acidity, fill=quality)) + geom_boxplot(color="red", aes(group=quality)) + facet_wrap(~quality, ncol = 3) + ylab("Fixed Acidity")+ stat_summary(fun.y = mean, geom="point", shape=5, size=1, color="green")
fixed.acidity_quality <- fixed.acidity_quality + theme(axis.text = element_text(angle=45, hjust = 1))
print(fixed.acidity_quality)


```

```{r echo=FALSE, Bivariate_Plots_Volatility.Acidity}

volatile.acidity_quality <- ggplot(rwdata, aes(x=quality, y=rwdata$volatile.acidity, fill=quality)) + geom_boxplot(color="orange", aes(group=quality)) + facet_wrap(~quality, ncol = 3) + ylab("Volatile Acidity")+ stat_summary(fun.y = mean, geom="point", shape=5, size=1, color="green")
volatile.acidity_quality <- volatile.acidity_quality + theme(axis.text = element_text(angle=45, hjust = 1))
print(volatile.acidity_quality)

```
```{r echo=FALSE, Bivariate_Plots_Citric.Acid}

citric.acid_quality <- ggplot(rwdata, aes(x=quality, y=rwdata$citric.acid, fill=quality)) + geom_boxplot(color="blue", aes(group=quality)) + facet_wrap(~quality, ncol = 3) + ylab("Citric Acid") + stat_summary(fun.y = mean, geom="point", shape=5, size=1, color="red")
citric.acid_quality <- citric.acid_quality + theme(axis.text = element_text(angle=45, hjust = 1))
print(citric.acid_quality)

```
> **Tip**: Make sure that you leave a blank line between the start / end of
each code block and the end / start of your Markdown text so that it is
formatted nicely in the knitted text. Note as well that text on consecutive
lines is treated as a single space. Make sure you have a blank line between
your paragraphs so that they too are formatted for easy readability.

```{r echo=FALSE, Bivariate_Plots_Residual.Sugar}
residual.sugar_quality <- ggplot(rwdata, aes(x=quality, y=rwdata$residual.sugar, fill=quality)) + geom_boxplot(color="green", aes(group=quality)) + facet_wrap(~quality, ncol = 3) + ylab("Residual Sugar") + stat_summary(fun.y = mean, geom="point", shape=5, size=1, color="red")
residual.sugar_quality <- residual.sugar_quality + theme(axis.text = element_text(angle=45, hjust = 1))
print(residual.sugar_quality)

```


```{r echo=FALSE, Bivariate_Plots_Chlorides}

chlorides_quality <- ggplot(rwdata, aes(x=quality, y=rwdata$chlorides, fill=quality)) + geom_boxplot(color="yellow", aes(group=quality)) + facet_wrap(~quality, ncol = 3) + ylab("Chlorides") + stat_summary(fun.y = mean, geom="point", shape=5, size=1, color="red")
chlorides_quality <- chlorides_quality + theme(axis.text = element_text(angle=45, hjust = 1))
print(chlorides_quality)

```


```{r echo=FALSE, Bivariate_Plots_Free.Sulphur.Dioxide}

free.sulfur.dioxide_quality <- ggplot(rwdata, aes(x=quality, y=rwdata$free.sulfur.dioxide, fill=quality)) + geom_boxplot(color="grey", aes(group=quality)) + facet_wrap(~quality, ncol = 3) + ylab("Free Sulfur Dioxide") + stat_summary(fun.y = mean, geom="point", shape=5, size=1, color="red")
free.sulfur.dioxide_quality <- free.sulfur.dioxide_quality + theme(axis.text = element_text(angle=45, hjust = 1))
print(free.sulfur.dioxide_quality)

```

```{r echo=FALSE, Bivariate_Plots_Total.Sulphur.Dioxide}

total.sulfur.dioxide_quality <- ggplot(rwdata, aes(x=quality, y=rwdata$total.sulfur.dioxide, fill=quality)) + geom_boxplot(color="magenta", aes(group=quality)) + facet_wrap(~quality, ncol = 3) + ylab("Total Sulfur Dioxide") + stat_summary(fun.y = mean, geom="point", shape=5, size=1, color="red")
total.sulfur.dioxide_quality <- total.sulfur.dioxide_quality + theme(axis.text = element_text(angle=45, hjust = 1))
print(total.sulfur.dioxide_quality)

```


```{r echo=FALSE, Bivariate_Plots_Density}

density_quality <- ggplot(rwdata, aes(x=quality, y=rwdata$density, fill=quality)) + geom_boxplot(color="brown", aes(group=quality)) + facet_wrap(~quality, ncol = 3) + ylab("Density") + stat_summary(fun.y = mean, geom="point", shape=5, size=1, color="red")
density_quality <- density_quality + theme(axis.text = element_text(angle=45, hjust = 1))
print(density_quality)

```

```{r echo=FALSE, Bivariate_Plots_PH}

pH_quality <- ggplot(rwdata, aes(x=quality, y=rwdata$pH, fill=quality)) + geom_boxplot(color="purple", aes(group=quality)) + facet_wrap(~quality, ncol = 3) + ylab("pH") + stat_summary(fun.y = mean, geom="point", shape=5, size=1, color="red")
pH_quality <- pH_quality + theme(axis.text = element_text(angle=45, hjust = 1))
print(pH_quality)

```


```{r echo=FALSE, Bivariate_Plots_Suplhates}

sulphates_quality <- ggplot(rwdata, aes(x=quality, y=rwdata$sulphates, fill=quality)) + geom_boxplot(color="black", aes(group=quality)) + facet_wrap(~quality, ncol = 3) + ylab("Sulphates") + stat_summary(fun.y = mean, geom="point", shape=5, size=1, color="red")
sulphates_quality <- sulphates_quality + theme(axis.text = element_text(angle=45, hjust = 1))
print(sulphates_quality)

```

```{r echo=FALSE, Bivariate_Plots_Alcohol}

alcohol_quality <- ggplot(rwdata, aes(x=quality, y=rwdata$alcohol, fill=quality)) + geom_boxplot(color="violet", aes(group=quality)) + facet_wrap(~quality, ncol = 3) + ylab("Alcohol") + stat_summary(fun.y = mean, geom="point", shape=5, size=1, color="red")
alcohol_quality <- alcohol_quality + theme(axis.text = element_text(angle=45, hjust = 1))
print(alcohol_quality)

```





```{r echo=FALSE, Bivariate_Plots}

```

# Bivariate Analysis


### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?


> Some interesting relationships I found in this dataset. When compared means of each feature \
with means of each feature with quality more than or equal to 8, I found that alcohol, total sulphur dioxide and \ 
free sulphur dioxde means are apart from quality means while all other are very close to quality means. It brings up \
a question whether these feature contribute more or less to the over all quality of wine.

> The boxplots above show a little more information beyond means. A new feature residual sugar has more dispersion \
around mean.


### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?


> The other features are distributed closely to the high quality features. Here is what I found 
Fixed Acidity is positively correlated with Density
Fixed Acidity is negatively correlated with PH. (expected!)
Volatile Acidity is positively correlated with PH (intereting!)
Volatile Acidity is negatively correlated with Citric Acid. It means something other than citric acid contributes to volatile acidity maybe?
Citric Acid and Fixed Acidity tend to behave similar.
Residual Sugar seem completely independent of all other features
Chlorides seem to go along well with Suplphates but seem to be independent of everything else.
Free Sulphur Dioxide seem slightly correlated with Total Sulphur Diaoxide but is independent of everything else.
Total Sulphur Dioxide is similar to Free Sulphur Dioxide. No relation seem exist with other factors.
Density seem to affect a few features. Correlates well with Fixed Acidity, Suplhates, Residual Sugar and Citric Acid. Seem to negatively correlate with Alcohol. This is interesting!
PH is mostly expected behavior. Negative to No correlation with Fixed Acidity and very tiny relation with Alcohol and Volatile Acidity.
Suplhates is interesting. It is going with no to slightly positively correlated with all other features.
Alcohol is not going along with any other feature except Density. It seem to negatively correlate with Density.

### What was the strongest relationship you found?

> Strongest relationship is Density. It is playing a role with several other factors although not effecting Quality directly. Wines with high quality has density really close to mean.


# Multivariate Plots Section



```{r echo=FALSE, Multivariate_Plots.Corr}

rwdata_corr <- select(rwdata, -(c(X)))
corrplot <- ggplot(melt(cor(rwdata_corr)), aes(x=Var1, y=Var2, fill=value)) + geom_tile() + theme(axis.text = element_text(angle=45, hjust=1))
print(corrplot)
```

```{r echo=FALSE, Multivariate_Plots.Model1}

infermodel <- lm(quality ~ pH + density + citric.acid + alcohol + total.sulfur.dioxide, data=rwdata_corr)
infermodel
summary(infermodel)
```

```{r echo=FALSE, Multivariate_Plots.InferPlotPH}
inferplotph <- ggplot(rwdata, aes(x=X)) + geom_point(aes(y=pH, color="blue")) + geom_point(aes(y=density, color="red")) + geom_point(aes(y=citric.acid, color="yellow")) + geom_point(aes(y=alcohol, color="green")) + geom_point(aes(y=total.sulfur.dioxide, color="orange")) + theme(axis.text = element_text(angle=45, hjust=1)) + scale_color_manual(name="All Factors Chosen", values=c("blue"="blue", "red"="red", "yellow"="yellow", "green"="green", "orange"="orange"), labels=c("pH", "Density", "Citric Acid", "Alcohol", "Total Sulfur Dioxide")) + ylab("Factors")
print(inferplotph)
```
```{r echo=FALSE, Multivariate_Plots.ABLine}

inferplotden <- ggplot(rwdata, aes(x=X)) + geom_point(aes(y=volatile.acidity, color="blue")) + geom_point(aes(y=density, color="red")) + geom_point(aes(y=residual.sugar, color="yellow")) + geom_point(aes(y=alcohol, color="green")) + theme(axis.text = element_text(angle=45, hjust=1)) + scale_color_manual(name="All Factors Chosen", values=c("blue"="blue", "red"="red", "yellow"="yellow", "green"="green", "orange"="orange"), labels=c("VolatileAcidity", "Density", "ResidualSugar", "Alcohol")) + ylab("Factors")
print(inferplotden)
```
```{r echo=FALSE, Multivariate_Plots}
generate_line <- function(data, intercept, ph, density, citric.acid, alcohol, total.sulfur.dioxide){
  linedata = data[,]
  linedata$line.point <- intercept + ph * data$pH + density * data$density + citric.acid * data$citric.acid + alcohol * data$alcohol + total.sulfur.dioxide * data$total.sulfur.dioxide
  return(linedata)
}

linedata <- generate_line(rwdata, 17.15, -0.455, -13.58, 0.6052, 0.3342, -0.00253)
regplot <- ggplot(linedata, aes(x = X)) + geom_point(aes(y=line.point, color="blue")) + geom_point(aes(y=quality, color="red")) + scale_color_manual(name="Regressed Line Points", values = c("blue"="blue", "red"="red"), labels=c("Regressed Line Point","Real Quality"))
print(regplot)
coeffs <- as.data.frame(coefficients(summary(infermodel))[1:6])
colnames(coeffs) <- c("Weights")
rownames(coeffs) <- rownames(coefficients(summary(infermodel)))
coeffplot <- ggplot(coeffs, aes(x=rownames(coeffs), y = Weights)) + geom_bar(aes(color="blue"), stat = "identity") + ylab("Coefficients") + xlab("Factors") + scale_color_manual(name="Coefficients", values =c("blue"="blue"), labels = c("Factors"))
print(coeffplot)
```


```{r echo=FALSE, Multivariate_Plots.Model2}

infermodel2 <- lm(quality ~ density + fixed.acidity + sulphates + residual.sugar, data=rwdata_corr)
infermodel2
summary(infermodel2)
coeffs2 <- as.data.frame(coefficients(summary(infermodel2))[1:5,c("Estimate")])
colnames(coeffs2) <- c("Coefficient")
coeffs2plot <- ggplot(coeffs2, aes(x = rownames(coeffs2), y = Coefficient)) + geom_bar(stat="identity") + ylab("Coefficients") + xlab("Factors")
print(coeffs2plot)
```


```{r echo=FALSE, Multivariate_Plots.Model3}

infermodel3 <- lm(quality ~ density + volatile.acidity + free.sulfur.dioxide + chlorides, data=rwdata_corr)
infermodel3
summary(infermodel3)
coeffs3 <- as.data.frame(coefficients(summary(infermodel3))[1:5,c("Estimate")])
colnames(coeffs3) <- c("Coefficient")
coeffs3plot <- ggplot(coeffs3, aes(x = rownames(coeffs3), y = Coefficient)) + geom_bar(stat="identity") + ylab("Coefficients") + xlab("Factors")
print(coeffs3plot)
```

```{r echo=FALSE, Multivariate_Plots.Model4}

infermodel4 <- lm(quality ~ density + alcohol, data=rwdata_corr)
infermodel4
summary(infermodel4)
coeffs4 <- as.data.frame(coefficients(summary(infermodel4))[1:3,c("Estimate")])
colnames(coeffs4) <- c("Coefficient")
coeffs4plot <- ggplot(coeffs4, aes(x = rownames(coeffs4), y = Coefficient)) + geom_bar(stat="identity") + ylab("Coefficients") + xlab("Factors")
print(coeffs4plot)
```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?


> I noticed a few relationships interesting to quote here. 
Total Sulphur Dioxide and Free Sulphur Dioxide overlap each other and we only need one of these.
Four other features interest me. Citric acid, density, ph and alcohol.
Citric acid controls fixed acidity and affects ph and alcohol and sulphates.
Density effects various other features as well
PH does not seem to effect other features positively but seem to negatively effect other features.
Alcohol is interesting as it seem to go along with other features but negatively effect density. 
From all these interactions it is hard to gather a combined effect of these features but they do present interesting behavior.


### Were there any interesting or surprising interactions between features?

> Yes, most surprising interaction is that none of the features directly effect quality that much. Along with this, these features do NOT interact with each other too much with a few expected negative correlations. This behavior indicates quality of wines cannot be simply explained by individual features even if these features show some interaction with each other.

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

> 


# Final Plots and Summary


### Plot One
```{r echo=FALSE, Plot_One}

rwdata_corr <- select(rwdata, -(c(X)))
corrplot <- ggplot(melt(cor(rwdata_corr)), aes(x=Var1, y=Var2, fill=value)) + geom_tile() + theme(axis.text = element_text(angle=45, hjust=1))
print(corrplot)
```

### Description One
> Plot one tells a lot about the dataset in a very small space. It compares every factor chosen with all other factors and shows them side by side. If you pick one row or one column, you can see how other factors correlate with the chosen factor (row/col we chose). The darker the color, the lower the correlation is, even negative. The brighter the color is the higher the correlation. The legend on the right tells range of values matching color. Lets take an example. Lets pick pH from 4th column from right. We will go from bottom to top. Bottom row and pH column is correlation with fixed.acidity. The color on that box is really dark indicating very low correlation between fixed.acidity and pH. Even negative. The next higher row in pH column is correlation with volatile.acidity. Comparing the color with the legend it indicates around 0.5. So on we can tell this relation ship for any factor.
> Note that the diagonal from top right to bottom left is the brightest. This make sense as each of this box is a correlation with itself which has to be 1.



### Plot Two
```{r echo=FALSE, Plot_Two}
inferplotden <- ggplot(rwdata, aes(x=X)) + geom_point(aes(y=volatile.acidity, color="blue")) + geom_point(aes(y=density, color="red")) + geom_point(aes(y=residual.sugar, color="yellow")) + geom_point(aes(y=alcohol, color="green")) + theme(axis.text = element_text(angle=45, hjust=1)) + scale_color_manual(name="All Factors Chosen", values=c("blue"="blue", "red"="red", "yellow"="yellow", "green"="green", "orange"="orange"), labels=c("VolatileAcidity", "Density", "ResidualSugar", "Alcohol")) + ylab("Factors")
print(inferplotden)
```

### Description Two

> I chose this plot because it shows the spread of factors in the distribution and how density dominates this spread. We have here three kinds of spreads. One very tightly distributed where most of the points are close to mean and none to little away from mean. These include Residual Sugar and Volatile Acidity. Second is data points close to mean and away but highly skewed. Meaning more data points on one side of the mean than other and hence breaking symmetry. Alcohol falls into this category. Third is most data points along the mean and the ones away from mean are more or less on both sides of the mean. Density falls into this category.

> It is very interesting to know how density is scatterred near normally. It raises curosity how it effects quality. That is what I explored in next plot. I like red wine personally and I keep trying new labels regularly and I noticed the change in wine quality with density. also noticed the change in color with density. Lets see how the next plot ties our intuition with facts.

### Plot Three
```{r echo=FALSE, Plot_Three}
infermodel4 <- lm(quality ~ density + alcohol, data=rwdata_corr)
infermodel4
summary(infermodel4)
coeffs4 <- as.data.frame(coefficients(summary(infermodel4))[1:3,c("Estimate")])
colnames(coeffs4) <- c("Coefficient")
coeffs4plot <- ggplot(coeffs4, aes(x = rownames(coeffs4), y = Coefficient)) + geom_bar(stat="identity") + ylab("Coefficients") + xlab("Factors")
print(coeffs4plot)
```

### Description Three
> I chose this plot to complete our exploration of above plot. This plot is a linear model showing effects of density and alcohol combined to quality of wine. The chart above is a plot of how coefficients of density and alcohol measure along with interception coefficient. The Console shows similar things in numbers and model output. As we show these coefficients, we understand that the coefficient of density is substantially larger than alcohol. This means that a small change in density substantially changes the quality of wine. As compared to density, alcohol coefficient is very small. Smalle coefficient means alcohol needs to change a lot for quality to change.
> Along with density and alcohol, we also notice that model is heavily biased with the intercept coefficient substantially large and negative. It tells us whatever density changes does to quality of wine, bias in model undoes it more or less. Overall, there is no good linear model to explain how these factors effect quality of wine!
------

# Reflection

> Data set explored here is already tidy so not much cleanup or mining to do here. That made exploration job little easier. The plan of action then is to explore each factor on its own, then see how they behave with each other and then notice how they behave with quality of wine. 

> For individual factors I plotted scatter and histogram along with basic descriptive statistics like mean etc... Scatter plot gives an initial intuition around distribution of a factor in the dataset. Histogram tunes it a little bit showing how close or far is the distribution to normal. Initially I tried to code each factor but then decided to plot in a loop. Looks more efficient that way. 

> Another step was to see how means of factors of wines with quality 8 or more compared with overall means of factors. Which factors mean were close the good wines means?

> Another step to continue explore how factors behave was to plot box plots filled by quality. I wanted to see how the distrbution looks like for each quality rating. Mins, Maxs, overall spreads, how far apart are some points from mean and the lower and higher quartile. Fitting box plots took some time and ggplot fiddling. Finally I got to a point where plots looks decent with some story to tell. It was interesting to find distribution of higher quality wines of most of our factors are pretty close to mean.

> Exploration would be incomplete without correlation plot of all factors with each other. This also took some learning and struggle. Along with learning it surprised me that none of the factors were highly correlated with quality. This is also the first time I found that density plays some important role.

> Correlation plot gave some important looking factors and I scatter plotted them together. Here also I noticed density playing an important role along with alcohol. It was not time to make models and see how all this intuition is following with numbers. I linear modelled some the combinations and finally density and alcohol together. I was surprised to see alcholol playing no role to effect quality. I was also surprised that none of the linear models clearly explained change in quality with change in factors. The R value is too little to give any major credibility to any of the models. One interesting find was to re assure what we intuited about density and what I see in real life as well! Good to know math works!

> It is clear that linear relationship does not play here. So next steps would be to apply non linear models like logistic regression. With more data, some deep learning methedology could also be utilized to learn how factors contribute to quality. SVMs can be used to cluster some factors and see if we can reduce some dimensionality and try other models like random forest.
